---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)

```
### Task 2: Load 1,000 rows of data with the ages of people from 18-80
```{r}
set.seed(42)

population <- tibble(
  age = sample (18:80, 1000, replace = TRUE)
)

population
```
### Task 3: Sample 100 rows of data from the dataset
```{r}
sampled_data <- population |>
  slice_sample(n = 100)

sampled_data
```
### Task 4: Calculate the mean age and standard deviation of the sample
```{r}
sample_stats <- sampled_data |> 
  summarise(
    mean_age = mean(age),
    sd_age = sd(age)
  )

sample_stats
```
### Task 5: Calculate the mean age and standard deviation of the population
```{r}

population_stats <- population |> 
  summarise(
    mean_age = mean(age),
    sd_age = sd(age)
  )

population_stats
```
### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

The sampled stats and the population stats are very similar, with the mean age only being .85 older in the population then it is in the sampled data and the standard deviation being .76 higher in the sampled data then it is in the population. With the sample being 50 it would make the data even worse, making the mean age over a year off and making the standard deviation smaller. With 25 people the mean is off by five years. The higher the sampled data the closer it will be to the population data. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
