---
title: "Line Chart Refresher: WMATA Ridership"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Quick Review: Making a Line Chart

Let's do a quick refresher on line charts using WMATA ridership data, since a line chart will be part of your homework. This is daily bus and rail data from January 2024 through February 2025 (you may recall it from the midterm). We'll aggregate it to weekly totals so the chart is easier to read, then make a line chart.

### Task 1: Load the tidyverse and the data

```{r}
library(tidyverse)
library(lubridate)

wmata <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")

head(wmata)
```

### Task 2: Make weekly totals

Daily data is noisy — weekends and holidays create big dips that make trends hard to see. Let's aggregate to weekly totals using `floor_date()` to group days into weeks, then `summarize()` to add them up.

```{r}
weekly_wmata <- wmata |>
  mutate(week = floor_date(date, "week")) |>
  group_by(week) |>
  summarize(
    weekly_total = sum(total),
    days = n()
  ) |>
  filter(days == 7)  # Keep only complete weeks

head(weekly_wmata)
```

### Task 3: Make a line chart

Create a line chart of weekly ridership over time.

```{r}
weekly_wmata |>
  ggplot(aes(x = week, y = weekly_total)) +
  geom_line(color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Weekly WMATA Ridership (2024-2025)",
    x = "Week",
    y = "Total Riders"
  ) +
  theme_minimal()
```

**Answer this question:** Describe the trend you see. Is ridership going up, down, or staying flat? What are the big dips, and what might explain them? What would you want to know before writing a story about Metro weekly ridership trends?

Looks like the ridership drops heavily around Janurary each year but besides that it was pretty still, with a couple large bumps like in June and in September and October. I think that the dips in late December/January can be explained by the Holidays. A lot of people are staying home or are out of town during the Holidays. Another reason could be less people commuting to their schools during winter break. I would want to know when there had been maintenance of any sort along with large events, like sports games or Holidays like the 4th of July.