---
title: "JOUR405: Statistics for Journalists"
subtitle: "Midterm Exam"
name: "Chase King"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

## Load the tidyverse

```{r}
library(tidyverse)
```

---

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible, the compliance score, and the grade.

### Task 1: Read the data (2 points)

```{r}
restaurants <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```

### Task 2: Calculate the mean, standard deviation, minimum, and maximum of compliance scores. Save the results in a new dataframe called `compliance_stats`. (2 points for each)

```{r}
compliance_stats <- restaurants |>
  summarize(
  compliance_mean = mean(compliance_score),
  compliance_sd = sd(compliance_score),
  compliance_min = min(compliance_score),
  compliance_max = max(compliance_score),
  )
```

### Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a descriptive title and label the x and y axes using `labs()`. (10 points)

```{r}
restaurants |>
  ggplot(compliance_stats, aes(x = compliance_score, y = grade))
  geom_histogram()
```

### Task 4: Write 2–3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What is newsworthy, and what would your next reporting step be? (5 points)

The standard deviation of the compliance stats was only 5.82 with the mean being 96.3. This means that the compliance stats of the restaurants were overall very similar with very few falling away from near perfect scores. 
---

## Part 2: Daily Transit Ridership — Z-Scores and Sampling (30 points)

Your editor wants you to identify unusual days in WMATA's daily ridership data that might be worth investigating as story leads. The dataset contains daily bus and rail totals. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv`

### Task 5: Load the WMATA ridership data. (2 points)

```{r}
WMATA_daily <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```

### Task 6: Using `summarize()`, calculate the mean and standard deviation for both bus and rail ridership. Save the results in a dataframe called `wmata_stats`. (5 points)

```{r}
wmata_stats <- WMATA_daily |>
  summarize(bus_mean = mean(bus), sd_bus = sd(bus), rail_mean = mean(rail), sd_rail = sd(rail))
```

### Task 7: Using `mutate()`, add two new columns to `wmata_daily`: `rail_zscore` (the z-score for rail ridership) and `bus_zscore` (the z-score for bus ridership). Save the result as `wmata_zscores`. (8 points)

```{r}
wmata_stats <- WMATA_daily  |>
  mutate(bus_mean = mean(bus), sd_bus = sd(bus), rail_mean = mean(rail), sd_rail = sd(rail))

wmata_zscores <- wmata_stats |>
  mutate(bus_zscore = mean(bus) / sd(bus), 
         rail_zscore = mean(rail) / sd(rail))
```

### Task 8: Filter `wmata_zscores` to show only days where the absolute value of `rail_zscore` is greater than 2 (unusually high or low rail ridership). Arrange by the absolute value of the z-score, with the most extreme first. How many unusual days are there? (8 points)

```{r}
unusual_days_wmata <- wmata_zscores |>
  filter(abs(rail_zscore)

```

### Task 9: Using `set.seed()`, take a random sample of 91 days from `wmata_daily` (roughly 25% of the data). Calculate the mean and standard deviation for bus and rail ridership in the sample. How do the sample statistics compare to the population statistics you calculated in Task 6? Write 1–2 sentences comparing the values and explaining what this suggests about random sampling. (4 points)

```{r}
set.seed(42)

sample_91 <- WMATA_daily |>
  sample_n(91)

sample_91 |>
  summarize(bus_mean = mean(bus), sd_bus = sd(bus), rail_mean = mean(rail), sd_rail = sd(rail))
```

  

### Task 10: Write 2–3 sentences: What do the unusual days in your z-score analysis suggest as story leads? (3 points)


---

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to find out which counties have the highest rates. The dataset contains car theft counts for Maryland counties in 2023 along with population figures. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv`

### Task 11: Load the car thefts data. (2 points)

```{r}
md_car_thefts <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

### Task 12: Using `mutate()`, add a column called `theft_rate` that calculates the car theft rate per 10,000 residents for each county. Save the result back to `md_car_thefts`. In 1–2 sentences, explain why a per-capita rate is more useful than raw counts for this comparison. (8 points)

```{r}
md_car_thefts  |>
  mutate(total_rate = thefts_2022 + thefts_2023,
         theft_rate = total_rate/population * 10000)
```
Per capita rate will always be the best way to compare cities/counties crime rates because it takes into account the size and gives a result per person. If a smaller city/county has a larger per capita crime rate that means you are more likely to fall victim to a crime then if it was a big city--which might have more overall crime--but a lower per capita rate.  

### Task 13: Calculate the median car theft rate and the total number of car thefts statewide. Show the counties that have rates above the median. (7 points)

```{r}
md_car_thefts  |>
  mutate(total_rate = thefts_2022 + thefts_2023,
         theft_rate = total_rate/population * 10000,
         median_rate = median(theft_rate),
         filter = theft_rate > median(theft_rate))

```

### Task 14: Write 2–3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What is the lede of a story based on your findings? (8 points)

The calculations reveal that the median rate of car rates is relatively low though there are some counties that have an extremely high car rates. I found out that Baltimore City, Prince George's County and St. Mary's County had the highest car theft rates.

Lede: Baltimore City is still way behind the rest of Maryland in car thefts.

---

## Part 4: Data Analysis Scenario (20 points)

A city parks department releases data showing that its 12 community swimming pools average 187 visitors per day during the summer season, with a standard deviation of 42 visitors. Your editor suspects that a handful of pools are dramatically underperforming and wants a story. You also learn that the department's figures are based on a random sample of 60 days per pool rather than a complete count.

Write 3–4 sentences for each of the following questions:

### Question A: Given that the mean is 187 and standard deviation is 42, which statistical measures would you prioritize using and why? (7 points)

I would use a zscore to see drastic differences in days that people are coming in or days that people avoid going to the pool. I would want to look at days of the weeks and see if a certain day has a lower zscore while other days have higher zscores. I think that using zscores will allow me to see whether or not any pools have consistently low attendances. 

### Question B: How would you use z-scores to identify which pools are genuinely underperforming, and why is that more useful than simply ranking pools by raw attendance? (7 points)

If there is a pool that has a lower zscore on average then I would be able to assume that they are under performing compared to other pools. I think it would also be important to look throughout the sample of days, see if the pools are receiving smaller z-scores as time goes on as it would mean people are opting not to come back. The higher a zscore ascends could also tell us the pools that are performing well with people which can be more important then ranking attendance because different pools are different sizes. 

### Question C: The department collected its attendance figures using a random sample of 60 days per pool rather than tracking every day. What questions would you ask about the sample before relying on these figures for a story? What concerns would you raise in your story? Use statistical terminology where appropriate. (6 points)

I would want to know what season it is in. If it is when the pools just open it will get less traffic then if it is in the middle of summer. I would want to know the tempature/ weather for each day because that can me a contributing factor to their success. I would want to know these things so that I could put some more context around any possible outliers, while also understanding why the sample looks like it does. If the sample was not taken in the middle of summer I would be scared that the data is incorrect as more people would be coming out to the pool during the middle of the summer. I want to make sure that all the pools were open on each day of the sample, because if not then it could ruin the zscores and I would need to account for it. 
---

### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your choices

Good luck!
