---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: "Chase King"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

In the class activity, you tested a school board member's claim using data from Montgomery County, Pa. — and found a model that didn't hold up. Now you'll run the same analysis using data from Montgomery County, Md. Your job is to evaluate whether the Maryland data tells a different story, and what that means for how you'd report on the claim.

### Loading the Data

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```

### Task 1: Examine the Data First (2 points)

Before you plot anything, look at the data — you know why by now.

```{r}
summary(schools$`FARMS Pct`)
schools |> arrange(desc(`FARMS Pct`)) |> head(10)
```

Are there any data quality issues like the ones you found in the Pennsylvania data? How does the distribution of FARMS percentages compare between the two counties? Note anything that concerns you or gives you confidence in the data.

There are not any schools that have a 100% FARMS percentage like there were in the Pennsylvania data, but there are some schools that have very high and very low FARMS percentages. The distribution is much more natural in the Maryland county.

### Task 2: Create a Scatterplot and Compare (5 points)

Replace the REPLACE_ME value with the correct variable name to create a scatterplot of FARMS percentage vs. math scores for Montgomery County, Md.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations",
       title = "Math Scores vs. FARMS Percentage, Montgomery County MD")
```

Describe what you see. Then compare this scatterplot specifically to the Pennsylvania one: what differences do you notice in the direction, strength, and spread of the relationship? The two counties share a name and are both suburban — what might explain why they look so different?

The scatterplot for the Pennsylvania county was very spread out and the line did not fully represent the data, but in the Maryland scatterplot the line represents the data much better as the data follows a stronger correlation. 

### Task 3: Linear Regression Model (5 points)

Replace the REPLACE_ME value with the correct variable name to run a linear regression model predicting math scores from FARMS percentage.

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)
summary(model)
```

Report the three key statistics: the p-value, the Adjusted R-squared, and the coefficient for FARMS Pct. Then interpret them together — not separately. What does the combination of these three numbers tell you about the relationship between poverty and math scores in Montgomery County, Md.? Is this model worth using? How does it compare to the Pennsylvania model?

The P-value is extremely low, being less then 0.00000000000000022 which is a good sign for the data. The adjusted R-squared is 0.6734 which is about ten times higher then the Pennsylvania data, which is another good sign. The coefficent for Farms Pct is also much lower then the Pennyslvania data which is another sign. Comparing the Maryland data and the Pennyslvania data, the Maryland data stands out as much stronger. 

### Task 4: Identifying Schools with the Biggest Residuals (6 points)

Replace the REPLACE_ME values with the correct variable names to calculate predicted values, residuals, and absolute residuals for each school. Then arrange the schools in descending order of absolute residuals.

```{r}
schools <- schools |>
  mutate(
    predicted = predict(model),
    residual = `% Met or Exceeded Math Expectations` - predicted,
    abs_residual = abs(residual)
  ) |>
  arrange(desc(abs_residual))

head(schools |> select(1, `FARMS Pct`, `% Met or Exceeded Math Expectations`, predicted, residual), 10)
```

Identify the school with the largest positive residual (doing much better than the model predicts) and the school with the largest negative residual (doing much worse). For each school, find its zip code and look it up on [CensusReporter.org](https://censusreporter.org).

For each of the two schools, answer:
- What does the demographic and economic profile of the zip code look like?
- Does anything in that profile help explain why the school is over- or under-performing the model?
- What would your first reporting question be to the principal of that school?

Fox Chapel Elementary has the largest positive residual and Germantown Elementary has the largest negative residual.

Fox Chapel Elementary: 
- $119,731 Median house income, 9.8% persons behind the poverty line, 8,783 households, 12% of population is 0-9 years old
- Nothing would really explain why there is such high scores from the census data.
- I would want to ask what they do to perpare students for their math tests 

Germantown Elementary:
- $115,022 Median House Income, 8.3% persons behind the poverty line, 22,662 households, 12% of population is 0-9 years old
- Nothing explains exactly why the residual test scores would be so low.
- I would want to ask what they've dont to address math test scores.

### Task 5: Evaluate the Model and Write the Pitch (7 points)

You've now run this analysis on two counties. Return to the school board member's claim: that elementary schools with fewer students on free and reduced-price meals do better on standardized math tests.

Write a short story pitch of 4-6 sentences that addresses all of the following: Does the Maryland data support or contradict the claim? How does it compare to what you found in Pennsylvania? What would you ask the school board member if you called them for comment? And what's the most interesting specific finding from your residuals analysis: which school would you want to visit first, and why?

It was foundthe data supported the clain that in schools with less students eligible for FARMS that there is better test scores. The data found much more correlation between FARMS percentage and test scores then the data in the Pennsylvania couny. I would like to ask the school board member what they are trying to do to address this and what steps need to be made. I would want to visit Harmony Hills Elementary, because they have low test scores correlated with a high FARMS percentage. I would also want to visit Fox Chapel Elementary school because they highest residual test scores and I would like to see what they are doing better then other schools.