---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2025.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

There are five different columns that describe the name, salary, position titles and more for the employees in the White House. There is eight salaries that really stick out as data that will cause problems for the analysis, and that is GOLDBERG RICHARD A., WITT PATRICK J., WHITE PAULA M., SACKS DAVID O., RUBIO MARCO A., MEANS CALLEY L., JOHNSON TRACY L.,  and DONALDSON DAVID P. because they all have a salary of 0 dollars. 

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |>
  summarize(
    mean_salary = mean(SALARY),
    median_salary = median(SALARY)
    )
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |>
  summarize(
    mean_salary = mean(SALARY),
    median_salary = median(SALARY),
    min_salary = min(SALARY),
    max_salary = max(SALARY)
    )
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The mean salary being higher than the median salary shows that there very high salaries that bring up the average over smaller salaries that show up more in the data. The minimum and maximum salaries show that there are outliers because both of them are over $100,000 away from the median. They also show that the salaries of higher positions can get really high, and that there are people in the White House's books that do not have listed salaries. 

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

A story idea could be about the eight people that are getting paid zero dollars in salary, and looking into what their roles are. I thought that "SACKS DAVID O." was interesting because his title was "WHITE HOUSE AI AND CRYPTO CZAR," which feels like a weird and interesting title. We can also look into the despirety in salaries and talk about how some assistants are only making $65 thousand dollars even though they are working for the most powerful person in the country.
