---
title: "HW9: Car Theft Confidence Intervals & Margin of Error"
name: REPLACE WITH YOUR NAME
---

Your editor wants to know what a "typical" month of car thefts looks like in Prince George's County. Not just the average — she wants to know how confident you can be in that number. Using monthly figures, confidence intervals and margin of error, you'll come up with an answer and decide whether the data supports a clear story. As with the previous homework, provide answers in this notebook when prompted and you will have to fill in values where you see REPLACE_ME. Get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# LOAD THE TIDYVERSE
```

### Task 2: Load the Prince George's county crime data

```{r}

# Read the data
car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/car_thefts_month.csv") |>
  arrange(month)  # Sort by date

# Basic inspection of the data
glimpse(car_thefts)

```

### Task 3: Let's make a chart and describe the results.

We'll make a line chart using geom_line() to try and illustrate any trend over time. What's your initial reaction to the change in car thefts? Is there a trend? If your editor asked you "how many car thefts happen in a typical month?", what would you say based on this chart alone?

It looks like the amount of car thefts is going down over time while also having spikes during the summer months. I wouldn't really be able to pinpoint a specific number of car thefts a month, but I could say that a lot of months have around 400 car thefts.

```{r}

car_thefts |>
ggplot(aes(x = month, y = total)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "darkblue", size = 3) +
  labs(
    title = "Monthly Car Thefts (July 2023 - February 2026)",
    subtitle = "Is there a trend over time?",
    x = "Month",
    y = "Number of Car Thefts"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold")
  )

```

### Task 4: Calculate the mean, standard error and margin of error

We've got a small number of months, so we need to quantify the uncertainty around our average. We'll calculate the standard error (which measures how much the sample mean could vary), and the margin of error (which gives us the plus/minus range). Replace the REPLACE_ME values based on what we did in class.

```{r}

theft_summary <- car_thefts |>
  summarize(
    mean_thefts = mean(total),
    sd_thefts = sd(total),
    months = n(),
    se_thefts = sd_thefts / sqrt(months),
    margin_error = qt(0.975, df = months - 1) * se_thefts
  )

# Display the results
theft_summary

```

**Answer this question:** Explain in your own words what the margin of error tells us here. If you were writing a story, how would you use this number? For example, how would you complete this sentence: "Prince George's County averages about ___ car thefts per month, plus or minus ___."

The margin of error tells us the amount that the mean could plausibly be wrong by. This means that the mean could be wrong by plus or minus 34.2 thefts. The margin of error being would be reason for less confidence in the data. Prince George's County averages about 373.6 car thefts per month plus or minus 34.2/

### Task 5: Build the confidence interval

Now let's turn the margin of error into a full confidence interval — the range where we're 95% confident the true average falls. Replace the REPLACE_ME values.

```{r}

theft_summary <- theft_summary |>
  mutate(
    ci_lower = mean_thefts - margin_error,
    ci_upper = mean_thefts + margin_error
  )

theft_summary |>
  select(mean_thefts, margin_error, ci_lower, ci_upper)

```

**Answer this question:** If a county official claimed "car thefts are down to about 300 per month," does your confidence interval support that claim? What about a claim of 350 per month? Explain how the confidence interval helps you evaluate each claim.

athe confidence interval would disagree with both claims that the thefts were down to 300 but it would not be able to shut down the claims that it dropped to below 350. The confidence interval gives an approximate value that represents a margin that the data could be in with 95% confidence. The confidence interval of a lower end of the thefts is only 341.4 thefts a month so the first claim would be wrong but the second claim couldn't be disproved.

### Task 6: Visualize the data with the confidence interval

Let's make a chart showing the data with the mean and the 95% confidence interval. Does this change anything about your answer from Task 3?

It does. I would now say that there were about 375 thefts were month, and in addition I would add that the confidence interval is from 341 to 409.

```{r}
# Make a chart with the mean and CI
car_thefts |>
ggplot(aes(x = month, y = total)) +
  geom_line(color = "steelblue", alpha = 0.7) +
  geom_point(size = 2) +
  geom_hline(yintercept = theft_summary$mean_thefts, linetype = "dashed", color = "red", size = 1) +
  geom_ribbon(aes(ymin = theft_summary$ci_lower, ymax = theft_summary$ci_upper),
              fill = "red", alpha = 0.2) +
  annotate("text", x = min(car_thefts$month),
           y = theft_summary$mean_thefts + 10,
           label = paste0("Mean: ", round(theft_summary$mean_thefts, 1)),
           hjust = 0, color = "red") +
  labs(
    title = "Monthly Car Thefts with 95% Confidence Interval for the Mean",
    x = "Month",
    y = "Number of Car Thefts"
  ) +
  theme_minimal()

```

### Task 7: The margin of error and sample size

In class we learned that MoE = 1/√n × 100 as a quick approximation for polls. That same principle applies here: more data means a smaller margin of error. Let's see what would happen if we had 40 months of data instead of the number we have, assuming the same mean and standard deviation. Replace the REPLACE_ME value with 40.

```{r}

hypothetical_summary <- tibble(
  mean_thefts = theft_summary$mean_thefts,
  sd_thefts = theft_summary$sd_thefts,
  months = 40,
  se_thefts = sd_thefts / sqrt(months),
  margin_error = qt(0.975, df = months - 1) * se_thefts,
  ci_lower = mean_thefts - margin_error,
  ci_upper = mean_thefts + margin_error
)

# Compare the two
comparison <- tibble(
  scenario = c("Actual data", "40 months"),
  months = c(theft_summary$months, hypothetical_summary$months),
  margin_error = c(theft_summary$margin_error, hypothetical_summary$margin_error),
  ci_lower = c(theft_summary$ci_lower, hypothetical_summary$ci_lower),
  ci_upper = c(theft_summary$ci_upper, hypothetical_summary$ci_upper)
)

comparison

```

**Answer this question:** How did the margin of error and confidence interval change? What does this tell you about the tradeoff between having more data and being able to make precise claims? If you were covering crime trends, how many months of data would you want before writing a definitive story?

The margin of error changed while the confidence interval shrunk from 341.3-409 to 345.2-405.9. The more data you have the more precise claims you can make, saying that the margin of error continues to shrink as you have more data. I would want around 40-50 months of data before writing a story.
### Task 8: Reporting with uncertainty

Based on everything you've calculated, answer the following:

1. Write a headline and a one-paragraph lede about car thefts in Prince George's County. Your text should properly reflect the uncertainty in the data — don't overstate or understate what the numbers show.

2. Your editor says "Just tell me if car thefts are going up or down." Using only the confidence interval (no significance testing yet), what's the most accurate answer you can give? What would you need to do next to answer that question more definitively?

Car thefts may be down in Prince George's County

1. Prince George's County has seen a decrease in car thefts in recent months, which many believe is the start of a new trend. The mean in car thefts over the last 32 months is 375.6, but the true numbers could between 341 to 406. It is common for months to differ in the total of thefts but the last six months have all had less than 341 thefts.

2. I would say that the thefts are going down sense there had been six straight months of lowering thefts that are all below the typical confidence interval of the mean but I would need a lot more months and month to month data analysis rather than data that clumps all 32 months together. 

When you are finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
