---
title: "final_project"
output: html_document
author: Chase King
---
Sources: 
2025 fall -> https://nacda.com/documents/2026/1/20/25.26DI_FinalFallConf.pdf 
2025 winter -> https://nacda.com/documents/2026/4/7/25.26DI_WinterConf_4.9.pdf 
Big ten athletic expenses -> https://www.extrapointsmb.com/p/oh-hey-lets-look-at-some-fy25-big-ten-athletic-budget-numbers, https://www.nytimes.com/athletic/7071017/2026/02/26/big-ten-financial-reports-2025/ 
possible tests:
standard deviation (mean of big ten rankings)
boxplot graph of standard deviation 
correlation between rankings and expenses 
scatterplot of correlation between rankings and expenses 
Linear Regression
Significance testing 


```{r}
library(tidyverse)
```

```{r}
college_athletic_rankings <- read_csv("https://raw.githubusercontent.com/chaseking731/jour405_spring26/refs/heads/main/project/Directors_Cup_Rankings.csv")
bigten_athletic_expenses <- read_csv("https://raw.githubusercontent.com/chaseking731/jour405_spring26/refs/heads/main/project/Big_Ten_Program_Budget_25Y(Sheet1).csv")
```

```{r}
college_athletic_rankings |>
  group_by(Conference) |>
  summarise(
    mean_total_points = mean(Total_Points),
    sd_total_points   = sd(Total_Points),
  )
```
This allows me to view the mean and standard deviation between different colleges and allows me to compare them in the future. I wanted to have data to compare to use to compare the big ten with other conferences. 

```{r}
combined_data <- college_athletic_rankings |>
  inner_join(bigten_athletic_expenses, by = c("Institution" = "College"))
```

I used AI to figure out gow to combine the two with "Institution and College being different. Link: https://chatgpt.com/share/69f02c63-e984-83ea-b995-0f9a2542d034 

```{r}
bigten_data <- combined_data |>
  filter(Conference == "Big Ten")
```

This just combines the data of only the big ten schools into a dataset for future use. 

```{r}
summary(bigten_data$`Total_Points`)
bigten_data |> arrange(desc(`Total_Points`)) |> head(10)
```

This shows some data from the bigten_data dataset. 


```{r}
bigten_data |>
  summarise(
    mean_total_points = mean(Total_Points),
    sd_total_points   = sd(Total_Points)
  )
```

This just gives me data about the mean from the big ten schools and the standard deviation.

```{r}
bigten_data |> summarize(correlation = cor(Total_Points, Expenses, method="pearson",))
```

This is not working with the data, it is saying that the data must be numerical. All of the data that is invovled with the code is numerical so I am not sure what it means. 


```{r}
bigten_data |>
  ggplot(aes(x = `Rank`, y = `Expenses`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Athletic Ranking", y = "Expenses",
       title = "Big Ten Schools Division 1 Athletic Ranking vs Athletic Expenses")
```

The code ran but it is completely unreadable. The expenses just show up as one datapoint. I think I need to set bounds for the X and Y axis but I am not sure how to do that. 

```{r}
bigten_data |>
  ggplot(aes(x = `Total_Points`, y = `Expenses`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Total Director Cup Points", y = "Expenses",
       title = "Big Ten Schools Division 1 Director Cup Points vs Athletic Expenses")
```
This is also unreadable because of the lack of bounds on the Y axis. But I think that it will be helpful once I have the bounds correct to see if there is a correlation between the total expenses and the total director cup points. 




Right now I have the data from the fall 2025-26 and winter 2025-26 NACDA Director's Cup Standings. I think that in the future when I am working with the data I might decide to go further back to collect more data. With winter_rankings and fall_rankings I will be able to filter out big ten teams and put together entire rankings to compare athletic programs. With the bigten_athletic_expenses I will be able to compare schools athletic expenses to their athletic program status as found with the winter_rankings and fall_rankings.